---
title: "Youth and Education Investment: A Cross-Country Analysis"
author: "Charvie Kukreja-24202861"
format:
  revealjs:
    theme: solarized
    slide-number: true
    toc: true
    incremental: true
    css: styles.css
    center: true   
fontsize: 22pt
mainfont: "Segoe UI"
---

# Introduction

This presentation explores economic, environmental, social, and development indicators across three countries:

-   **USA**
-   **Luxembourg**
-   **India**

We use `data.table` for efficient data handling and `ggplot2` for visualization.

```{r}
#| include: false
# Load required packages
library(data.table)
library(ggplot2)
library(skimr)
```

# Narrative and Objective

We investigate the relationship between **education investment** and **youth development**, focusing on:

-   Youth not in education, employment or training

-   Gender parity in school enrollment

-   Tertiary education attainment

-   Adolescent fertility rate

# Data Overview

Source: \[Humanitarian Data Exchange - World Bank Indicators\]

Indicators from:

-   India (IND)

-   United States (USA)

-   Luxembourg (LUX)

Years: varies per indicator (mostly 1990–2022)

# Loading and Merging Data

```{r}
# Load files
ind <- fread("indicators_ind.csv")
usa <- fread("indicators_usa.csv")
lux <- fread("indicators_lux.csv")

# Merge
dt <- rbindlist(list(ind, usa, lux))
dt
```

# Filter Selected Indicators

```{r}
indicators <- c(
  "Share of youth not in education, employment or training, total (% of youth population)",
  "School enrollment, primary and secondary (gross), gender parity index (GPI)",
  "Government expenditure on education, total (% of GDP)",
  "Educational attainment, at least completed post-secondary, population 25+, total (%) (cumulative)",
  "Adolescent fertility rate (births per 1,000 women ages 15-19)"
)

dt_sel <- dt[`Indicator Name` %in% indicators]
dt_sel
```

# Variable Summary

```{r}
skim(dt_sel)
```

# Youth not in Education/training

```{r}
neet <- dt_sel[`Indicator Name` == indicators[1]]

# Better styled plot
ggplot(neet, aes(x = Year, y = Value, color = `Country Name`, group = `Country Name`)) +
  geom_line(size = 1.3) +
  geom_point(size = 2.5) +
  scale_color_brewer(palette = "Set1") +
  labs(
    title = "Youth Not in Education, Employment or Training (NEET)",
    subtitle = "Trends across India, Luxembourg, and United States",
    y = "% of Youth (15–24)",
    x = "Year",
    color = "Country"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    plot.subtitle = element_text(size = 13),
    axis.title = element_text(face = "bold"),
    legend.position = "bottom"
  )
```

-   India shows the **highest NEET rate** and a **rising trend**, highlighting youth disengagement.

-   Luxembourg maintains consistently low levels.

-   The US exhibits a moderate but more stable pattern.

# Gender Parity in Enrollment

```{r}
gpi <- dt_sel[`Indicator Name` == indicators[2]]
gpi[, `:=`(Value = as.numeric(Value), Year = as.integer(Year))]

ggplot(gpi, aes(x = Year, y = Value, color = `Country Name`, group = `Country Name`)) +
  geom_line(size = 1.3) +
  geom_point(size = 2.5) +
  scale_color_brewer(palette = "Set1") +
  labs(
    title = "Gender Parity Index (Primary & Secondary)",
    subtitle = "School Enrollment Parity (F/M Ratio)",
    y = "Index (0–1)",
    x = "Year",
    color = "Country"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    plot.subtitle = element_text(size = 13),
    axis.title = element_text(face = "bold"),
    legend.position = "bottom"
  )
```

-   India and Luxembourg are near parity or slightly favor females.

-   USA maintains consistent balance.

-   Gender disparity appears largely addressed.

# Education Spending (% GDP)

```{r}
ed_spending <- dt_sel[`Indicator Name` == indicators[3]]
ed_spending[, `:=`(Value = as.numeric(Value), Year = as.integer(Year))]

ggplot(ed_spending, aes(x = Year, y = Value, color = `Country Name`, group = `Country Name`)) +
  geom_line(size = 1.3) +
  geom_point(size = 2.5) +
  scale_color_brewer(palette = "Set1") +
  labs(
    title = "Government Education Expenditure (% of GDP)",
    subtitle = "Investment in Public Education by Country",
    y = "% of GDP",
    x = "Year",
    color = "Country"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    plot.subtitle = element_text(size = 13),
    axis.title = element_text(face = "bold"),
    legend.position = "bottom"
  )
```

-   India invests 4%(approx) of GDP in education, the **lowest** among the three.

-   Luxembourg shows the **highest and most stable investment**.

-   U.S. exhibits moderate but consistent funding.

# Tertiary Attainment (Age 25+)

```{r}
tertiary <- dt_sel[`Indicator Name` == indicators[4]]
tertiary[, `:=`(Value = as.numeric(Value), Year = as.integer(Year))]

ggplot(tertiary, aes(x = Year, y = Value, color = `Country Name`, group = `Country Name`)) +
  geom_line(size = 1.3) +
  geom_point(size = 2.5) +
  scale_color_brewer(palette = "Set1") +
  labs(
    title = "Tertiary Education Attainment (25+)",
    subtitle = "Share of Population with Post-secondary Education",
    y = "% of Adults (25+)",
    x = "Year",
    color = "Country"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    plot.subtitle = element_text(size = 13),
    axis.title = element_text(face = "bold"),
    legend.position = "bottom"
  )
```

-   Luxembourg and the USA lead in tertiary attainment.

-   India lags behind and shows **declining** attainment rates.

-   Indicates need for sustained investment in higher education.

# Adolescent Fertility Rate

```{r}
fertility <- dt_sel[`Indicator Name` == indicators[5]]
fertility[, `:=`(Value = as.numeric(Value), Year = as.integer(Year))]

ggplot(fertility, aes(x = Year, y = Value, color = `Country Name`, group = `Country Name`)) +
  geom_line(size = 1.3) +
  geom_point(size = 2.5) +
  scale_color_brewer(palette = "Set1") +
  labs(
    title = "Adolescent Fertility Rate",
    subtitle = "Births per 1,000 Women Ages 15–19",
    y = "Births per 1,000 Women",
    x = "Year",
    color = "Country"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    plot.subtitle = element_text(size = 13),
    axis.title = element_text(face = "bold"),
    legend.position = "bottom"
  )
```

-   India has seen a **rise** in adolescent fertility, indicating gaps in health education.

-   Luxembourg and the U.S. show **lower and decreasing** rates.

-   Higher fertility correlates with lower educational attainment.

# Summary Tables(Average Values)

```{r}
# Ensure 'Value' is numeric
dt_sel[, Value := as.numeric(Value)]

# Compute summary table
sum_tbl <- dt_sel[
  , .(Mean_Value = mean(Value, na.rm = TRUE)), 
  keyby = .(`Country Name`, `Indicator Name`)
]
print(sum_tbl)

```

#Decision : Cross Country Trends

-   **India**: Higher NEET rate and fertility, lower tertiary education.

-   **Luxembourg**: High parity index and spending.

-   **USA**: Steady investment, but large variation in outcomes.

#Interpretation:

-   **Education spending** and **gender parity** are strong predictors of youth inclusion.

-   Countries with low fertility often have higher tertiary attainment.

# Efficient data.table Code Example

```{r}
# Max NEET year by country
neet[, .SD[which.max(Value)], keyby = `Country Name`]
```

# Personalisation Notes

-   Used **custom LaTeX header** for fonts and spacing.

-   Fonts: Fira Sans.

-   Beamer theme: compact with clean lines.

# Final Thoughts

-   Investment in education correlates strongly with youth development.

-   Gender-focused policies improve parity and outcomes.

# References

-   Humanitarian Data Exchange: https://data.humdata.org/

-   World Bank Education & Development Indicators
